\textit{[něco jako?]} Generating lambda term individuals using forgetful A*

### 2013-10-22 tex/evogp.tex
evogp tex

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\Real}{\mathbb{R}}

\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}

@@

\title{\textit{[něco jako?]} 
Generating lambda term individuals using undisciplined A*}
\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
\institute{Matfyz ...}
\maketitle

@@

\subsubsection{Reductions}
\subsubsection{Tree representations}
%\subsubsection{}

@@

@@

\textit{--- nezapomenout popsat $\Gamma,x:\tau$ notaci
 a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
 tzn je nová...}

\subsection{\GMS}

@@

\textit{- Unifished term je term který může obsahovat jako list uzel tvaru 
$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\
-Vysvětlit unfinished leaf.\\}

\section{Our approach}
\subsection{Introduction}
úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek

@@

Essential data structure of our term generating algorithm 
is priority queue of unfinished terms. 
Priority of an unfinished term is given by 
the number of unfinished leafs in that term.
Initially, the queue contains only one unfinished term; 
$\ul{\tau}{\Gamma}$.\\

At each step, the term $M$ with the lowest number of 
unfinished leafs is pulled from the queue.
According to the actual number of those leafs one of
the following actions is performed.
\begin{enumerate}
 \item If the term has no unfinished leaf (i.e., it is a finished
 term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
 result collection of generated terms.   
 \item Otherwise, \textit{successors} of the unfinished term $M$ are
       filtered out by \textit{search strategy} $S$ and
       those successors that outlast the filtration 
       are inserted into the queue.
\end{enumerate}

@@

\textit{- pač je to generovany v lnf, neboli beta-eta na -1 nf kde 
eta na -1 je eta expanze tak je chytrý transformovat 
to do beta-eta nf. To stačí opakovanou eta redukcí, protože beta normálnost
se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
}

@@

\textit{Zmínil bych jí tu ale jen hodně rychle...}

>>>>

\usepackage[ampersand]{easylist}
\usepackage{color}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\C}{\mathbb{C}\xspace}
\newcommand{\Real}{\mathbb{R}}

\newenvironment{todo}
{ ~\\[0.5em]
  {\color{red}\textbf{TODO}}
  \begin{easylist}[itemize]}
{ \end{easylist}
  ~}

\newcommand{\Lpr}{\Lambda^\prime}
\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}

@@

\title{\textit{[něco jako?]} 
Generating lambda term individuals using 
%undisciplined 
forgetful
A*}
\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
\institute{Matfyz ...}
\maketitle

@@

Here we  %formally 
describe programming language, 
in which we generate individual programs --- so called \lterms.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\textit{Function application} and 
\textit{$\lambda$-abstraction} are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}.

\begin{todo}
 & asi říct o zkratkách notací
\end{todo}

\subsubsection{Reductions}
(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
\subsubsection{Tree representations}
(... zmiňovat ?)

@@

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).
After putting those two pieces 
(\textit{\lterms} and \textit{types}) together 
we will get system called \textit{simply typed $\lambda$-calculus}.

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}~

\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$.

\begin{todo}
 & asi říct o zkratce notace
\end{todo}

@@

\begin{definition} 
~
\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 %(or \textit{basis}) 
 is set of declarations with distinct variables as subjects.
\end{enumerate}~
\end{definition}

\begin{todo}
   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
   & popsat $\Gamma,x:\tau$ notaci
 a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
 tzn je nová...
  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
\end{todo}

\subsection{Simply typed $\lambda$-calculus}

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}~
\end{definition}

...Our goal is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.

\begin{todo}
    & From this definition we can derive following rules...

	& citovat něco, asi barendrechta 
	(on to tam ale nějak moc nedokazuje) 
	případně mojí diplomku (tam to dokazuju)
\end{todo}

@@

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	
Let $\T$ be set of types.
Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
Then $\Lpr$ is set of 
\textit{unfinished  \lterms} defined inductively as follows.	
\begin{align*}
\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
\textit{~~~~~~~~(Unfinished leaf)}\\
x   \in V \cup C  &\then x     \in \Lpr \\
M,N \in \Lpr   &\then (M~N) \in \Lpr 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\begin{todo}
   & Vysvětlit k čemu je to dobrý.
\end{todo}

\section{Our approach}
\subsection{Introduction}

\begin{todo}
 & úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek
\end{todo}

@@

Essential data structure of our term generating algorithm 
is priority queue of unfinished terms. 
Priority of an unfinished term is given by its size.
Initially, the queue contains only one unfinished term; 
$\ul{\tau}{\Gamma}$.\\

At each step, the term $M$ with the smallest size
is pulled from the queue.
According to the actual number of those leafs one of
the following actions is performed.
\begin{enumerate}
 \item If the term $M$ has no unfinished leaf (i.e., it is a finished
 term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
 result collection of generated terms.   
 \item Otherwise, \textit{successors} of the unfinished term $M$ are
       filtered out by \textit{search strategy} $S$ and
       those successors that outlast the filtration 
       are inserted into the queue.
\end{enumerate}

@@

\textbf{(Otázka zda vůbec zminovat?)}

\textit{- pač je to generovany v lnf, neboli \beenf kde 
$\eta^{-1}$ je $\eta$-expanze tak je chytrý transformovat 
to do \benf. To stačí opakovanou $\eta$-redukcí, protože beta normálnost
se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
}

@@

\textit{Zmínil bych jí tu ale jen hodně rychle..., nutno zmínit
pokud budem uvádět even parity problém }
